---
output: github_document
bibliography: bibliography.bib
nocite: |
       @tahmed, @bahadori, @Dranchuk1973, @Kamyab2010, @Rakap2016, @Al-Anazi2011, @Elsharkawy2001, @Mohamadi-Baghmolaei2015, @Kumar2004, @Almeida2014, @Elsharkawy2001a, @Standing1942, @Kareem2016, @Kassem1990, @Fatoorehchi2014, @Hall1973
biblio-style: apalike
link-citations: yes
---

[![Rdoc](http://www.rdocumentation.org/badges/version/zFactor)](http://www.rdocumentation.org/packages/zFactor)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# zFactor
Computational tools for chemical, petrochemical and petroleum engineers. Calculates the deviation between different correlations for gas compressibility against the Standing-Katz chart.

## Motivation
For the development of vertical lift performance (VLP) curves and other petroleum engineering calculations, it is necessary to calculate properties of hydrocarbons in mixtures, liquid and gaseous phases.  Compressibility is one of these hydrocarbon properties. It is important to count with the apropriate correlation and know its range of applicability in order to calculate other properties dependant of `z`.

There are several compressibility correlations. In this package, few of them, or the most used are being evaluated. There has been extensive studies on compressibility correlations. This evaluation is different in the sense that provides a graphical view of the range of applicability as well statistical measures of the errors.

This analysis focus on sweet hydrocarbon gases. Compressibility correlations for sour gases are little bit more complicated since they show significant deviations from the curves in the Standing-Katz charts. Effects of CO2, H2S and N2 have to accounted for. In a future release of `zFactor` correlations for sour hydrocarbons gases will be covered.

## Correlations used in this study
The correlations that are implemented in R for `zFactor` are:

* Beggs and Brill (BB)
* Hall and Yarborough (HY)
* Dranchuk and Abou-Kassem (DAK)
* Dranchuk, Purvis and Robnson (DPR)
* A correlation by Shell Oil Company (SH)
* A correlation developed with Artificial Neural Networks (Ann10) by Kamyab et al. See footnotes

Note. The Shell correlation was found cited in two books and two papers but the original paper or authors could not be identified.


## Installation
I recommend installing from GitHub using devtools. CRAN release cycles only allow updates every one to two months.

You can install the latest version of `zFactor` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("f0nzie/zFactor")
```


Or from `CRAN`:

```{r eval=FALSE}
install.packages("zFactor")
```

## Usage

```{r, message = FALSE}
library(zFactor)

# get `z` values from the Standing-Katz chart
tpr <- c(1.2, 1.3, 1.5, 2.0, 3.0) 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 
getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr, 
                                 pprRange = "lp")

# calculate `z` using the Beggs-Brill correlation
z.BeggsBrill(pres.pr = 1.5, temp.pr = 2.0)

# calculate `z` using the Hall-Yarborough correlation
z.HallYarborough(pres.pr = 1.5, temp.pr = 2.0)

# calculate `z` using the correlation Dranchuk-AbousKassem
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.05, 1.1, 1.7, 2) 
z.DranchukAbuKassem(ppr, tpr)

# We do the same with the Dranchuk-Purvis-Robinson correlation
# but we change the values of the isotherms `Tpr`
tpr <- c(1.4, 1.4, 1.6, 1.7, 1.8) 
z.DranchukPurvisRobinson(pres.pr = ppr, temp.pr = tpr)

# now we use the relative undocumented Shell correlation
z.Shell(ppr, tpr)

# and finally the correlation Kamyab et al that uses Artificial Neural Networks
z.Ann10(ppr, tpr)

```

## Comparative Analysis

## Range of Applicability


```{r}
library(zFactor)

zFactor:::z.plot.range("HY", interval = "coarse")
zFactor:::z.plot.range("BB")

zFactor:::z.plot.range("DAK")

zFactor:::z.plot.range("DPR")

zFactor:::z.plot.range("SH")
zFactor:::z.plot.range("SH", interval = "fine")

zFactor:::z.plot.range("N10")
```


## What you can do with `zFactor`

* Find `z` with any of the correlations provided
* Get values from the Standing-Katz chart at any of the isotherms limited by the pseudo-reduced pressures (`Ppr`) digitized
* Find what isotherms or pseudo-reduced temperatures (`Tpr`) are available from Standing and Katz chart
* Find what pseudo-reduced pressure points are available
* Plot any of the Standing-Katz isotherms to view `z` in graphical form
* Plot a tile chart to show the range of applicability of the correlations
* Get a statistics table when comparing any of the correlations to the Standing-Katz chart
* Get a matrix of ``z` values calculated from any correlation where the rows are the pseudo-reduced temperatures and the columns the pseudo-reduced pressures


## Vignettes
The vignettes contain examples on the use and analysis of the various correlations.

* StandingKatz_chart.Rmd
* Beggs-Brill.Rmd
* Hall-Yarborough.Rmd
* Dranchuk-AbouKassem.Rmd
* Dranchuk-Purvis-Robinson.Rmd
* shell.Rmd
* ANN.Rmd

## Tests
There are tests for the correlations under tests/testthat.

## References
The following books and papers were consulted during the development of this package:







