---
output: github_document
---

[![Rdoc](http://www.rdocumentation.org/badges/version/zFactor)](http://www.rdocumentation.org/packages/zFactor)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```


```{r echo=FALSE}
library(knitcitations)
knitcitations::cleanbib()
bib_file <- system.file("doc", "bibliography.bib", package = "zFactor")
bib <- read.bibtex(bib_file)
```


# zFactor
Computational tools for chemical, petrochemical and petroleum engineers. Calculates the deviation between different correlations for gas compressibility compared to the Standing-Katz chart.

## Motivation
For the development of vertical lift performance (VLP) curves and other petroleum engineering calculations, it is necessary to calculate properties of hydrocarbons in mixtures, liquid and gaseous phases.  Compressibility is one of these hydrocarbon properties. It is important to count with the apropriate correlation and know its range of applicability in order to calculate other properties dependant of `z`.

There are several compressibility correlations. In this package, few of them, or the most used are being evaluated. There has been extensive studies on compressibility correlations. This evaluation is different in the sense that provides a graphical view of the range of applicability as well statistical measures of the errors.

This analysis focus on sweet hydrocarbon gases. Compressibility correlations for sour gases are little bit more complicated since they show significant deviations from the curves in the Standing-Katz charts. Effects of CO2, H2S and N2 have to accounted for. In a future release of `zFactor` correlations for sour hydrocarbons gases will be covered.

## Correlations used in this study
The correlations that are implemented in R for the package **zFactor** are:

* Beggs and Brill (BB) ``r citep(bib[c("Azizi2010")])`` 

* Hall and Yarborough (HY) ``r citep(bib[c("Hall1973")])`` 

* Dranchuk and Abou-Kassem (DAK) ``r citep(bib[c("Dranchuk1975")])`` 

* Dranchuk, Purvis and Robinson (DPR) ``r citep(bib[c("Dranchuk1973")])`` 

* A correlation by Shell Oil Company (SH) ``r citep(bib[c("Kumar2004")])``, ``r citep(bib["bahadori"])``, ``r citep(bib["Almeida2014"])``, ``r citep(bib["Al-Anazi2011"])``, ``r citep(bib["Azizi2010"])``, ``r citep(bib["Mohamadi-Baghmolaei2015"])``. The Shell correlation was found cited in two books and several papers (including the equation and constants), but the original paper or authors could not be identified. 

* A correlation developed with Artificial Neural Networks (Ann10) by Kamyab et al. ``r citep(bib[c("Kamyab2010")])``


## Installation
I recommend installing from GitHub using devtools, that way you get the latest and greatest version. CRAN release cycles allows updates every one to two months.

You can install the latest version of `zFactor` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("f0nzie/zFactor")
```


Or, if your prefer, from `CRAN`:

```{r eval=FALSE}
install.packages("zFactor")
```

## Usage

```{r, message = FALSE}
library(zFactor)

# get `z` values from the Standing-Katz chart
tpr <- c(1.2, 1.3, 1.5, 2.0, 3.0) 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 
getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr, 
                                 pprRange = "lp")

# calculate `z` using the Beggs-Brill correlation
z.BeggsBrill(pres.pr = 1.5, temp.pr = 2.0)

# calculate `z` using the Hall-Yarborough correlation
z.HallYarborough(pres.pr = 1.5, temp.pr = 2.0)

# calculate `z` using the correlation Dranchuk-AbousKassem
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.05, 1.1, 1.7, 2) 
z.DranchukAbuKassem(ppr, tpr)

# We do the same with the Dranchuk-Purvis-Robinson correlation
# but we change the values of the isotherms `Tpr`
tpr <- c(1.4, 1.4, 1.6, 1.7, 1.8) 
z.DranchukPurvisRobinson(pres.pr = ppr, temp.pr = tpr)

# now we use the relative undocumented Shell correlation
z.Shell(ppr, tpr)

# and finally the correlation Kamyab et al that uses Artificial Neural Networks
z.Ann10(ppr, tpr)

```


## Range of Applicability
How to interpret the colors? We use the Mean Absolute Percentage Error or `MAPE` to visualize how close the correlation follow the experimental values of the Standing-Katz chart.

* `Blue`: the MAPE is zero or near zero
* `Yellow`: the MAPE is around 10 percent.
* 'Red`: the MAPE has reached 25% or more
* `Pink`: the mean absolute percentage error is above or way above 25%.

You can see for yourself which correlation is more stable at different ranges of pseudo-reduce pressures and temperatures.

```{r}
library(zFactor)

zFactor:::z.plot.range("HY",  interval = "fine")
zFactor:::z.plot.range("BB",  interval = "fine")
zFactor:::z.plot.range("DAK", interval = "fine")
zFactor:::z.plot.range("DPR", interval = "fine")
zFactor:::z.plot.range("SH",  interval = "fine")
zFactor:::z.plot.range("SH",  interval = "fine")
zFactor:::z.plot.range("N10", interval = "fine")
```


## Comparative Analysis
The comparative analysis shows tables with different error measurements:

    MSE:   Mean Squared Error
    RMSE:  Root Mean Sqyared Error
    RSS:   Residual sum of Squares
    RMSLE: Root Mean Squared Logarithmic Error. Penalizes understimation.
    MAPE:  Mean Absolute Percentage Error = AARE
    MPE:   Mean Percentage error = ARE
    MAE:   Mean Absolute Error

```{r}
# Example of errors table

```


## What you can do with `zFactor`

* Find `z` with any of the correlations provided
* Get values from the Standing-Katz chart at any of the isotherms limited by the pseudo-reduced pressures (`Ppr`) digitized
* Find what isotherms or pseudo-reduced temperatures (`Tpr`) are available from Standing and Katz chart
* Find what pseudo-reduced pressure points are available
* Plot any of the Standing-Katz isotherms to view `z` in graphical form
* Plot a tile chart to show the range of applicability of the correlations
* Get a statistics table when comparing any of the correlations to the Standing-Katz chart
* Get a matrix of ``z` values calculated from any correlation where the rows are the pseudo-reduced temperatures and the columns the pseudo-reduced pressures


## Vignettes
The vignettes contain examples on the use and analysis of the various correlations.

* StandingKatz_chart.Rmd
* Beggs-Brill.Rmd
* Hall-Yarborough.Rmd
* Dranchuk-AbouKassem.Rmd
* Dranchuk-Purvis-Robinson.Rmd
* shell.Rmd
* ANN.Rmd

## Tests
There are tests for the correlations under tests/testthat.

## References
The following books and papers were consulted during the development of this package:

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
```

```{r results="asis", echo=FALSE}
bibliography()
```










