---
output: github_document
---

[![Rdoc](http://www.rdocumentation.org/badges/version/zFactor)](http://www.rdocumentation.org/packages/zFactor)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# zFactor
Computational tools for chemical and petroleum engineers. Calculates the error between different correlations for gas compressibility.

## Motivation

The goal of rODE is to explore R and its S4 classes and the differences with Java and Python classes while exploring simulations by solving differential equations.


The correlations that are implemented in R for `zFactor` are:

* Beggs and Brill (BB)
* Hall and Yarborough (HY)
* Dranchuk and Abou-Kassem (DAK)
* Dranchuk, Purvis and Robnson (DPR)
* A correlation by Shell Oil Company (SH)
* A correlation developed with Artificial Neural Networks (Ann10) by Kamyab et al

Note. The Shell correlation was found cited in two books and two papers but the original paper or authors could not be found.


## Installation
You can install the latest version of `zFactor` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("f0nzie/zFactor")
```


Or from `CRAN`:

```{r eval=FALSE}
install.packages("zFactor")
```

## Usage

```{r, message = FALSE}
library(zFactor)

# get `z` values from the Standing-Katz chart
tpr <- c(1.2, 1.3, 1.5, 2.0, 3.0) 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 
getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr, 
                                 pprRange = "lp")

# calculate `z` using the Beggs-Brill correlation
z.BeggsBrill(pres.pr = 1.5, temp.pr = 2.0)

# calculate `z` using the Hall-Yarborough correlation
z.HallYarborough(pres.pr = 1.5, temp.pr = 2.0)

# calculate `z` using the correlation Dranchuk-AbousKassem
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.05, 1.1, 1.7, 2) 
z.DranchukAbuKassem(ppr, tpr)

# We do the same with the Dranchuk-Purvis-Robinson correlation
# but we change the values of the isotherms `Tpr`
tpr <- c(1.4, 1.4, 1.6, 1.7, 1.8) 
z.DranchukPurvisRobinson(pres.pr = ppr, temp.pr = tpr)

# now we use the relative undocumented Shell correlation
z.Shell(ppr, tpr)

# and finally the correlation Kamyab et al that uses Artificial Neural Networks
z.Ann10(ppr, tpr)

```

## Comparative Analysis

```{r}

```


## Vignettes
The vignettes contain examples on the use and analysis of the various correlations.

* StandingKatz_chart.Rmd
* Beggs-Brill.Rmd
* Hall-Yarborough.Rmd
* Dranchuk-AbouKassem.Rmd
* Dranchuk-Purvis-Robinson.Rmd
* shell.Rmd
* ANN.Rmd

## Tests
There are tests for the correlations under tests/testthat.




