---
output: html_document
bibliography: bibliography.bib
biblio-style: apalike
link-citations: yes
---

[![Rdoc](http://www.rdocumentation.org/badges/version/zFactor)](http://www.rdocumentation.org/packages/zFactor)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# zFactor
Computational tools for chemical and petroleum engineers. Calculates the error between different correlations for gas compressibility.

## Motivation



The correlations that are implemented in R for `zFactor` are:

* Beggs and Brill (BB)
* Hall and Yarborough (HY)
* Dranchuk and Abou-Kassem (DAK)
* Dranchuk, Purvis and Robnson (DPR)
* A correlation by Shell Oil Company (SH)
* A correlation developed with Artificial Neural Networks (Ann10) by Kamyab et al [-@Kamyab2010]

Note. The Shell correlation was found cited in two books and two papers but the original paper or authors could not be found.


## Installation
I recommend installing from GitHub using devtools. CRAN release cycles only allow updates every one to two months.

You can install the latest version of `zFactor` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("f0nzie/zFactor")
```


Or from `CRAN`:

```{r eval=FALSE}
install.packages("zFactor")
```

## Usage

```{r, message = FALSE}
library(zFactor)

# get `z` values from the Standing-Katz chart
tpr <- c(1.2, 1.3, 1.5, 2.0, 3.0) 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 
getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr, 
                                 pprRange = "lp")

# calculate `z` using the Beggs-Brill correlation
z.BeggsBrill(pres.pr = 1.5, temp.pr = 2.0)

# calculate `z` using the Hall-Yarborough correlation
z.HallYarborough(pres.pr = 1.5, temp.pr = 2.0)

# calculate `z` using the correlation Dranchuk-AbousKassem
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.05, 1.1, 1.7, 2) 
z.DranchukAbuKassem(ppr, tpr)

# We do the same with the Dranchuk-Purvis-Robinson correlation
# but we change the values of the isotherms `Tpr`
tpr <- c(1.4, 1.4, 1.6, 1.7, 1.8) 
z.DranchukPurvisRobinson(pres.pr = ppr, temp.pr = tpr)

# now we use the relative undocumented Shell correlation
z.Shell(ppr, tpr)

# and finally the correlation Kamyab et al that uses Artificial Neural Networks
z.Ann10(ppr, tpr)

```

## Comparative Analysis

## Range of Applicability


```{r}
library(zFactor)

zFactor:::z.plot.range("HY", interval = "coarse")
zFactor:::z.plot.range("BB")

zFactor:::z.plot.range("DAK")

zFactor:::z.plot.range("DPR")

zFactor:::z.plot.range("SH")
zFactor:::z.plot.range("SH", interval = "fine")

zFactor:::z.plot.range("N10")
```


## What you can do with `zFactor`

* Find `z` with any of the correlations provided
* Get values from the Standing-Katz chart at any of the isotherms limited by the pseudo-reduced pressures (`Ppr`) digitized
* Find what isotherms or pseudo-reduced temperatures (`Tpr`) are available from Standing and Katz chart
* Find what pseudo-reduced pressure points are available
* Plot any of the Standing-Katz isotherms to view `z` in graphical form
* Plot a tile chart to show the range of applicability of the correlations
* Get a statistics table when comparing any of the correlations to the Standing-Katz chart
* Get a matrix of ``z` values calculated from any correlation where the rows are the pseudo-reduced temperatures and the columns the pseudo-reduced pressures


## Vignettes
The vignettes contain examples on the use and analysis of the various correlations.

* StandingKatz_chart.Rmd
* Beggs-Brill.Rmd
* Hall-Yarborough.Rmd
* Dranchuk-AbouKassem.Rmd
* Dranchuk-Purvis-Robinson.Rmd
* shell.Rmd
* ANN.Rmd

## Tests
There are tests for the correlations under tests/testthat.

## References
The following books and papers were consulted during the development of this package:







