---
title: "Dranchuk-AbouKassem correlation"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dranchuk-AbouKassem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=F, error=T, message=F, warning=F}
knitr::opts_chunk$set(echo=T, comment=NA, error=T, warning=F, message = F, fig.align = 'center', results="hold")
```


## How do we find the limits of accuracy in the `DAK` correlation


## Get `z` at selected `Ppr` and `Tpr`

```{r}
# get a z value
library(zFactor)

z.DranchukAbuKassem(pres.pr = 1.5, temp.pr = 2.0)
# HY = 0.9580002
```

From the Standing-Katz chart we obtain a digitized point at the same `Ppr` and `Tpr`:

```{r}
# get a z value from the SK chart at the same Ppr and Tpr
library(zFactor)

tpr_vec <- c(2.0)
getStandingKatzMatrix(tpr_vector = tpr_vec, 
                      pprRange = "lp")[1, "1.5"]
```

It looks pretty good.


## Get `z` at selected `Ppr` and `Tpr`

```{r}
library(zFactor)

z.DranchukAbuKassem(pres.pr = 1.5, temp.pr = 1.1)
```

From the `Standing-Katz` chart we obtain a digitized point:

```{r}
library(zFactor)

tpr_vec <- c(1.1)
getStandingKatzMatrix(tpr_vector = tpr_vec, 
                      pprRange = "lp")[1, "1.5"]
```

> At lower `Tpr` we can see that there is some difference we see a  difference between the values of z from the DAK calculation and the value read from the Standing-Katz chart.


## Get values of `z` for several `Ppr` and `Tpr`

```{r}
# test HY with 1st-derivative using the values from paper 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.3, 1.5, 1.7, 2) 

dak <- z.DranchukAbuKassem(ppr, tpr)
print(dak)
 
# From Hall-Yarborough
#    0.5       1.5       2.5       3.5       4.5       5.5       6.5
# 1.3 0.9176300 0.7534433 0.6399020 0.6323003 0.6881127 0.7651710 0.8493794
# 1.5 0.9496855 0.8581232 0.7924067 0.7687902 0.7868071 0.8316848 0.8906351
# 1.7 0.9682547 0.9134862 0.8756412 0.8605668 0.8694525 0.8978885 0.9396353
# 2   0.9838234 0.9580002 0.9426939 0.9396286 0.9490995 0.9697839 0.9994317
```


With the same `ppr` and `tpr` vectors, we do the same for the Standing-Katz chart:

```{r}
library(zFactor)

sk <- getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr)
sk
```

Subtract and find the difference:

```{r}
err <- round((sk - dak) / sk * 100, 2)
err
```

## Error by Ppr and by PPr

```{r}
print(colSums(err))
```

```{r}
print(rowSums(err))
```

## Analyze the error for smaller values of `Tpr`

```{r}
library(zFactor)

tpr2 <- c(1.05, 1.1) 
ppr2 <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 


sk2 <- getStandingKatzMatrix(ppr_vector = ppr2, tpr_vector = tpr2, pprRange = "lp")
sk2
```


We do the same with the DAK correlation:

```{r}
# calculate z values at lower values of Tpr
library(zFactor)

dak2 <- z.DranchukAbuKassem(pres.pr = ppr2, temp.pr = tpr2) 
 
print(dak2)
```

Calculate the error:

```{r}
err2 <- round((sk2 - dak2) / sk2 * 100, 2)
err2
```

Transposing the matrix with `Tpr` as columns and `Ppr` as rows:

```{r}
t_err2 <- t(err2)
t_err2
```


A statistical summary by `Tpr` curve:

```{r}
sum_t_err2 <- summary(t_err2)
sum_t_err2
```

We can see that the errors in `z` with `DAK` are lower than `Hall-Yarborough` with a `r sum_t_err2[1,1]`% and `r sum_t_err2[6,1]`% for `Tpr=1.05`, and a `r sum_t_err2[1,2]`% and `r sum_t_err2[6,2]`% for `Tpr=1.10`. They are significant errors.


## Prepare to plot SK vs DAK correlation

```{r}
library(zFactor)
library(tibble)

tpr2 <- c(1.05, 1.1, 1.2, 1.3) 
ppr2 <- c(0.5, 1.0, 1.5, 2, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0, 6.5) 

sk_dak_2 <- createTidyFromMatrix(ppr2, tpr2, correlation = "DAK")
as.tibble(sk_dak_2)
```

The error bars represent the difference between the calculated values by the `Dranchuk-AbouKassem` corrrelation and teh values of `z` read from the Standing-Katz chart.

```{r}
library(ggplot2)

p <- ggplot(sk_dak_2, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-dif, ymax=z.calc+dif), width=.4,
                  position=position_dodge(0.05))
print(p)
```

We can see with Dranchuk-AbouKassem there are still errors or differences between the values in the Standing-Katz chart and the values obtained from the correlation but they are not so bad as in the HY correlation.

## Analysis at the lowest `Tpr`

```{r}
sk_dak_3 <- sk_dak_2[sk_dak_2$Tpr==1.05,]
sk_dak_3
```


```{r}
library(zFactor)

p <- ggplot(sk_dak_3, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-dif, ymax=z.calc+dif), width=.2,
                  position=position_dodge(0.05))
print(p)
```

```{r}
summary(sk_dak_3)
```


## Analyzing performance of the `DAK` correlation for all the `Tpr` curves

```{r}
library(ggplot2)
library(tibble)

# get all `lp` Tpr curves
tpr_all <- getCurvesDigitized(pprRange = "lp")
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
sk_corr_all <- createTidyFromMatrix(ppr, tpr_all, correlation = "DAK")
as.tibble(sk_corr_all)


p <- ggplot(sk_corr_all, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-dif, ymax=z.calc+dif), width=.4,
                  position=position_dodge(0.05))
print(p)
```

```{r}
# MSE: Mean Squared Error
# RMSE: Root Mean Sqyared Error
# RSS: residual sum of square
# ARE:  Average Relative Error, %
# AARE: Average Absolute Relative Error, %
library(dplyr)
grouped <- group_by(sk_corr_all, Tpr, Ppr)
smry_tpr_ppr <- summarise(grouped, 
          RMSE= sqrt(mean((z.chart-z.calc)^2)), 
          MSE = sum((z.calc - z.chart)^2) / n(), 
          RSS = sum((z.calc - z.chart)^2),
          ARE = sum((z.calc - z.chart) / z.chart) * 100 / n(),
          AARE = sum( abs((z.calc - z.chart) / z.chart)) * 100 / n()
          )

ggplot(smry_tpr_ppr, aes(Ppr, Tpr)) + 
    geom_tile(data=smry_tpr_ppr, aes(fill=AARE), color="white") +
    scale_fill_gradient2(low="blue", high="red", mid="yellow", na.value = "pink",
                         midpoint=12.5, limit=c(0, 25), name="AARE") + 
    theme(axis.text.x = element_text(angle=45, vjust=1, size=11, hjust=1)) + 
    coord_equal() +
    ggtitle("Dranchuk-AbouKassem", subtitle = "DAK")
```


> The greatest errors are localized in two of the Tpr curves: 1.05 and barely at 1.1


## Looking numerically at the errors in `DAK` vs `SK` chart

```{r}
# get all `lp` Tpr curves
tpr <- getCurvesDigitized(pprRange = "lp")
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 

# calculate HY for the given Tpr
all_dak <-z.DranchukAbuKassem(pres.pr = ppr, temp.pr = tpr)
cat("Calculated from correlation \n")
print(all_dak) 

cat("\nStanding-Katz chart\n")
all_sk <- getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr)
all_sk

# find the error
cat("\n Errors in percentage \n")
all_err <- round((all_sk - all_dak) / all_sk * 100, 2)  # in percentage
all_err

cat("\n Errors in Ppr\n")
summary(all_err)

# for the transposed matrix
cat("\n Errors for the transposed matrix: Tpr \n")
summary(t(all_err))
```


