---
title: "Hall-Yarborough correlation"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hall-Yarborough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=F, error=T, message=F, warning=F}
knitr::opts_chunk$set(echo=T, comment=NA, error=T, warning=F, message = F, fig.align = 'center')
```


## How do we find the limits of accuracy in the HY correlation



## Get `z` at selected `Ppr` and `Tpr`

```{r}
# get a z value using HY
library(zFactor)

z.HallYarborough(pres.pr = 1.5, temp.pr = 2.0)
```

From the Standing-Katz chart we obtain a digitized point:

```{r}
# get a z value from the SK chart at the same Ppr and Tpr
library(zFactor)

tpr_vec <- c(2.0)
getStandingKatzMatrix(tpr_vector = tpr_vec, 
                      pprRange = "lp")[1, "1.5"]
```



## Get `z` at selected `Ppr` and `Tpr`

```{r}
library(zFactor)

z.HallYarborough(pres.pr = 1.5, temp.pr = 1.1)
```

From the Standing-Katz chart we obtain a digitized point:

```{r}
library(zFactor)

tpr_vec <- c(1.1)
getStandingKatzMatrix(tpr_vector = tpr_vec, 
                      pprRange = "lp")[1, "1.5"]
```

> We perceive a noticeable difference between the values of z from the HY calculation and the value read from the Standing-Katz chart.


## Get values of `z` for several `Ppr` and `Tpr`

```{r}
# test HY with 1st-derivative using the values from paper 
 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.3, 1.5, 1.7, 2) 
 
hy <- sapply(ppr, function(x)  
    sapply(tpr, function(y) z.HallYarborough(pres.pr = x, temp.pr = y))) 
 
rownames(hy) <- tpr 
colnames(hy) <- ppr 
print(hy) 
```

With the same ppr and tpr vector, we do the same for the Standing-Katz chart:

```{r}
library(zFactor)

sk <- getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr)
sk
```

Subtract and find the difference:

```{r}
err <- round((sk - hy) / sk * 100, 2)
err
```

## Error by Ppr and by PPr

```{r}
print(colSums(err))
```

```{r}
print(rowSums(err))
```

## Analyze the error for smaller values of `Tpr`

```{r}
library(zFactor)

tpr2 <- c(1.05, 1.1) 
ppr2 <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 


sk2 <- getStandingKatzMatrix(ppr_vector = ppr2, tpr_vector = tpr2, pprRange = "lp")
sk2
```

```{r}
# find what common values of PPr are available in the chart
getStandingKatzMatrix(tpr_vector = c(1.05))
```

We do the same with the HY correlation:

```{r}
# calculate z values at lower values of Tpr
library(zFactor)

hy2 <- sapply(ppr2, function(x)  
    sapply(tpr2, function(y) z.HallYarborough(pres.pr = x, temp.pr = y))) 
 
rownames(hy2) <- tpr2
colnames(hy2) <- ppr2
print(hy2)
```

```{r}
err2 <- round((sk2 - hy2) / sk2 * 100, 2)
err2
```

We can see that using Hall-Yarborough correlation shows a very high error at values of `Tpr` lower or equal than `1.1` being `Tpr=1.05` the worst curve to extract z values from.

```{r}
t_err2 <- t(err2)
t_err2
```


```{r}
sum_t_err2 <- summary(t_err2)
sum_t_err2
```

We can see that the errors in `z` are considerable with a `r sum_t_err2[1,1]`% and `r sum_t_err2[6,1]`% for `Tpr=1.05`, and a `r sum_t_err2[1,2]`% and `r sum_t_err2[6,2]`% for `Tpr=1.10`

## Prepare to plot SK vs HY correlation

```{r}
library(zFactor)

tpr2 <- c(1.05, 1.1, 1.2, 1.3) 
ppr2 <- c(0.5, 1.0, 1.5, 2, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0, 6.5) 

sk_hy_2 <- createTidyFromMatrix(ppr2, tpr2, correlation = "HY")
sk_hy_2
```


```{r}
library(ggplot2)

p <- ggplot(sk_hy_2, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-dif, ymax=z.calc+dif), width=.4,
                  position=position_dodge(0.05))
print(p)
```


## Analyzing the error for all the `Tpr` curves

```{r}
library(ggplot2)

# get all `lp` Tpr curves
tpr_all <- getCurvesDigitized(pprRange = "lp")
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
sk_hy_all <- createTidyFromMatrix(ppr, tpr_all, correlation = "HY")
sk_hy_all


p <- ggplot(sk_hy_all, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-dif, ymax=z.calc+dif), width=.4,
                  position=position_dodge(0.05))
print(p)
```

## Looking numerically at the errors

```{r}
# get all `lp` Tpr curves
tpr <- getCurvesDigitized(pprRange = "lp")
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 

# calculate HY for the given Tpr
all_hy <- sapply(ppr, function(x)  
    sapply(tpr, function(y) z.HallYarborough(pres.pr = x, temp.pr = y))) 
rownames(all_hy) <- tpr 
colnames(all_hy) <- ppr 
cat("Calculated Hall-Yarborough\n")
print(all_hy) 

cat("\nStanding-Katz chart\n")
all_sk <- getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr)
all_sk

# find the error
cat("\n Errors in percentage \n")
all_err <- round((all_sk - all_hy) / all_sk * 100, 2)  # in percentage
all_err

cat("\n Errors in Ppr\n")
summary(all_err)

# for the transposed matrix
cat("\n Errors for the transposed matrix: Tpr \n")
summary(t(all_err))
```

