---
title: "Hall-Yarborough"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hall-Yarborough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## How do we find the limits of accuracy in the HY correlation



## Get `z` at selected `Ppr` and `Tpr`

```{r}
library(zFactor)

z.HallYarborough(pres.pr = 1.5, temp.pr = 2.0)
```

From the Standing-Katz chart we obtain a digitized point:

```{r}
library(zFactor)

tpr_vec <- c(2.0)
getStandingKatzMatrix(tpr_vector = tpr_vec, 
                      pprRange = "lp")[1, "1.5"]
```



## Get `z` at selected `Ppr` and `Tpr`

```{r}
library(zFactor)

z.HallYarborough(pres.pr = 1.5, temp.pr = 1.1)
```

From the Standing-Katz chart we obtain a digitized point:

```{r}
library(zFactor)

tpr_vec <- c(1.1)
getStandingKatzMatrix(tpr_vector = tpr_vec, 
                      pprRange = "lp")[1, "1.5"]
```

> We perceve noticeable difference between the values of z from the HY calculation and the value read from the Standing-Katz chart.


## Get values of `z` for several `Ppr` and `Tpr`

```{r}
# test HY with 1st-derivative using the values from paper 
 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.3, 1.5, 1.7, 2) 
 
hy <- sapply(ppr, function(x)  
    sapply(tpr, function(y) z.HallYarborough(pres.pr = x, temp.pr = y))) 
 
rownames(hy) <- tpr 
colnames(hy) <- ppr 
print(hy) 
```

With the same ppr and tpr vector, we do the same for the Standing-Katz chart:

```{r}
library(zFactor)

sk <- getStandingKatzMatrix(ppr_vector = ppr_vec, tpr_vector = tpr_vec)
sk
```

Subtract and find the difference:

```{r}
err <- (sk - hy) / sk * 100
err
```

## Error by Ppr and by PPr

```{r}
print(colSums(err))
```

```{r}
print(rowSums(err))
```

## Analyze the error for smaller values of Tpr

```{r}
library(zFactor)


ppr2 <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 
tpr2 <- c(1.05, 1.1) 

sk2 <- getStandingKatzMatrix(ppr_vector = ppr2, tpr_vector = tpr2, pprRange = "lp")
sk2
```

```{r}
# find what common values of PPr are available in the chart
getStandingKatzMatrix(tpr_vector = c(1.05))
```

We do the same with the HY correlation:

```{r}
library(zFactor)

ppr2 <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 
tpr2 <- c(1.05, 1.1) 

hy2 <- sapply(ppr2, function(x)  
    sapply(tpr2, function(y) z.HallYarborough(pres.pr = x, temp.pr = y))) 
 
rownames(hy2) <- tpr2
colnames(hy2) <- ppr2
print(hy2)
```

```{r}
err2 <- (sk2 - hy2) / sk2 * 100
err2
```

We can see that using Hall-Yarborough correlation shows a very high error at values of Tpr lower or equal than 1.1 being Tpr=1.05 the worst curve to extract z values from.
