---
title: "Hall-Yarborough correlation"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{Hall-Yarborough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=F, error=T, message=F, warning=F}
knitr::opts_chunk$set(echo=T, comment=NA, error=T, warning=F, message = F, fig.align = 'center', results="hold")
```


## The Hall-Yarborough correlation
Kenneth Hall and Lyman Yarborough used the hard-sphere equation as the basis for the equation of state. They tested the correlation with 12 reservoir gas reservoir systems up to `Ppr` as high as 20.5. The Standing-Katz chart only extends to `Ppr=15`. At that moment the Standing-Katz chart had 30 years of existance. See [@Hall1973].

## Get `z` at selected `Ppr` and `Tpr=2.0`

Use the the corelation to calculate `z`. From the Standing-Katz chart we obtain a digitized point at the given `Tpr` and `Ppr`.

```{r}
# get a z value using HY
library(zFactor)

ppr <- 1.5
tpr <- 2.0

z.calc <- z.HallYarborough(pres.pr = ppr, temp.pr = tpr)

# get a z value from the SK chart at the same Ppr and Tpr
z.chart <- getStandingKatzMatrix(tpr_vector = tpr, 
                      pprRange = "lp")[1, as.character(ppr)]

# calculate the APE (Average Percentage Error)
ape <- abs((z.calc - z.chart) / z.chart) * 100

df <- as.data.frame(list(Ppr = ppr,  z.calc =z.calc, z.chart = z.chart, ape=ape))
rownames(df) <- tpr
df
```


## Get `z` at selected `Ppr` and `Tpr = 1.1`
From the Standing-Katz chart we obtain a digitized point:

```{r}
library(zFactor)
ppr <- 1.5
tpr <- 1.1

z.calc <- z.HallYarborough(pres.pr = ppr, temp.pr = tpr)

# From the Standing-Katz chart we obtain a digitized point:
z.chart <- getStandingKatzMatrix(tpr_vector = tpr, 
                      pprRange = "lp")[1, as.character(ppr)]

# calculate the APE
ape <- abs((z.calc - z.chart) / z.chart) * 100

df <- as.data.frame(list(Ppr = ppr,  z.calc =z.calc, z.chart = z.chart, ape=ape))
rownames(df) <- tpr
df
```

> We see here a noticeable difference between the values of `z` from the `HY` correlation and the value read from the Standing-Katz chart. This is expected at these low pressures and temperatures. This area is a challenge for all of the correlations as well.


## Get values of `z` for combinations of `Ppr` and `Tpr`
In this example we provide vectors instead of a single point.
With the same `ppr` and `tpr` vectors that we use for the correlation, we do the same for the `Standing-Katz` chart. We want to compare both and find the `absolute percentage error`.

```{r}
library(zFactor)

ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.05, 1.1, 1.7, 2) 

# calculate using the correlation
z.calc <- z.HallYarborough(ppr, tpr)

# With the same ppr and tpr vector, we do the same for the Standing-Katz chart
z.chart <- getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr)
ape <- abs((z.calc - z.chart) / z.chart) * 100

# calculate the APE
cat("z.correlation \n"); print(z.calc)
cat("\n z.chart \n"); print(z.chart)
cat("\n APE \n"); print(ape)
```

> You can see errors of 22.48% and 12.09% in the isotherm `Tpr=1.05` at `Ppr=1.5 and 2.5, respectively. Other errors, greater than one, can also be found at the isotherm `Tpr=1.1`. Then, the rest of the curves are fine.

## Analyze the error at the `isotherms`
We apply the function `summary` over the transpose of the matrix. Applying the transpose will alow us to see the statistics at each of the isotherms.

```{r}
sum_t_ape <- summary(t(ape))
sum_t_ape
```

> We see that the errors in `z` are considerable with a `r sum_t_ape[1,1]`% and `r sum_t_ape[6,1]`% for `Tpr=1.05`, and a `r sum_t_ape[1,2]`% and `r sum_t_ape[6,2]`% for `Tpr=1.10`. 

The Hall-Yarborough correlation shows a very high error at values of `Tpr` lower or equal than `1.1`, being `Tpr=1.05` the worst curve to calculate z values from. Keep that in mind. We will explore later a comparative **tile** chart where we confirm these early calculations.



## Analyze the error for greater values of `Tpr`
Let's see the numbers now for higher values of `Tpr` at various values of `Ppr`.

```{r}
library(zFactor)
# enter vectors for Tpr and Ppr
tpr2 <- c(1.2, 1.3, 1.5, 2.0, 3.0) 
ppr2 <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 
# get z values from the SK chart
z.chart <- getStandingKatzMatrix(ppr_vector = ppr2, tpr_vector = tpr2, pprRange = "lp")

# We do the same with the HY correlation:
# calculate z values at lower values of Tpr
z.calc <- z.HallYarborough(pres.pr = ppr2, temp.pr = tpr2) 
ape <- abs((z.calc - z.chart) / z.chart) * 100

# calculate the APE
cat("z.correlation \n"); print(z.calc)
cat("\n z.chart \n"); print(z.chart)
cat("\n APE \n"); print(ape)
```

> At `Tpr` above or equal to 1.2 the `HY` correlation behaves very well.


## Analyze the error at the `isotherms`
Now, let's apply  the function `summary` over the transpose of the matrix to observe the error of the correlation at each isotherm.

```{r}
sum_t_ape <- summary(t(ape))
sum_t_ape
```

> In all, the error of `z` at these isotherms is less than 0.82%. **Pretty good**.



## Prepare to plot `SK` chart values vs `HY` correlation
Now, we will be plotting the difference between the `z` values in the Standing-Katz and the `z` values calculated by Hall-Yarborough. For the moment let's not pay attention to the numerical error but to the `error bars` in the plot (the orange bars).


```{r}
library(zFactor)
library(tibble)
library(ggplot2)

tpr2 <- c(1.05, 1.1, 1.2, 1.3)
ppr2 <- c(0.5, 1.0, 1.5, 2, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0, 6.5)

sk_corr_2 <- createTidyFromMatrix(ppr2, tpr2, correlation = "HY")
as.tibble(sk_corr_2)

p <- ggplot(sk_corr_2, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-abs(dif), ymax=z.calc+abs(dif)), width=.4,
                  position=position_dodge(0.05))
print(p)
```

> As we were expecting, the errors can be found on Tpr=1.05 and Tpr=1.1


## Analyzing the error for all the `Tpr` curves
In this last example, we compare the values of `z` at all the isotherms.
We use the function `getCurvesDigitized` in the `zfactor` package to obtain all the isotherms or `Tpr` curves in the Standing-Katz chart that have been digitized. The next function `createTidyFromMatrix` calculates `z` using the correlation and prepares a tidy dataset ready to plot.

```{r}
library(ggplot2)
library(tibble)

# get all `lp` Tpr curves
tpr_all <- getStandingKatzTpr(pprRange = "lp")
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
sk_corr_all <- createTidyFromMatrix(ppr, tpr_all, correlation = "HY")
as.tibble(sk_corr_all)

p <- ggplot(sk_corr_all, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-dif, ymax=z.calc+dif), width=.4,
                  position=position_dodge(0.05))
print(p)
```

> As we saw before, we confirm that the greatest errors are localized in two of the `Tpr` curves: at 1.05 and 1.1.


## Range of applicability of the correlation
What we will see here is the distribution of the Mean Average Percentage Error or MAPE accross all the `Tpr` isotherms and `Ppr` grids. Remember from the `README`: red and yellow are bad.
 
```{r}
# MSE: Mean Squared Error
# RMSE: Root Mean Sqyared Error
# RSS: residual sum of square
# RMSLE: Root Mean Squared Logarithmic Error. Penalizes understimation.
# MAPE: Mean Absolute Percentage Error = AARE
# MPE: Mean Percentage error = ARE
# MAE: Mean Absolute Error
library(dplyr)
grouped <- group_by(sk_corr_all, Tpr, Ppr)
smry_tpr_ppr <- summarise(grouped, 
          RMSE= sqrt(mean((z.chart-z.calc)^2)), 
          MPE = sum((z.calc - z.chart) / z.chart) * 100 / n(),
          MAPE = sum(abs((z.calc - z.chart) / z.chart)) * 100 / n(), 
          MSE = sum((z.calc - z.chart)^2) / n(), 
          RSS = sum((z.calc - z.chart)^2),
          MAE = sum(abs(z.calc - z.chart)) / n(),
          RMLSE = sqrt(1/n()*sum((log(z.calc +1)-log(z.chart +1))^2))
          )

ggplot(smry_tpr_ppr, aes(Ppr, Tpr)) + 
    geom_tile(data=smry_tpr_ppr, aes(fill=MAPE), color="white") +
    scale_fill_gradient2(low="blue", high="red", mid="yellow", na.value = "pink",
                         midpoint=12.5, limit=c(0, 25), name="MAPE") + 
    theme(axis.text.x = element_text(angle=45, vjust=1, size=11, hjust=1)) + 
    coord_equal() +
    ggtitle("Hall-Yarborough", subtitle = "HY")
```




## Plotting the `Tpr` and `Ppr` values that show more error
This is just plotting the couple of isotherms where we see the largest errors.

```{r}
library(dplyr)

sk_corr_all %>%
    filter(Tpr %in% c("1.05", "1.1")) %>%
    ggplot(aes(x = z.chart, y=z.calc, group = Tpr, color = Tpr)) +
    geom_point(size = 3) +
    geom_line(aes(x = z.chart, y = z.chart), color = "black") +
    facet_grid(. ~ Tpr) +
    geom_errorbar(aes(ymin=z.calc-abs(dif), ymax=z.calc+abs(dif)),
                  position=position_dodge(0.05))
```


## Looking numerically at the errors
Finally, the dataframe with the calculated errors between the `z` from the correlation and the `z` read from the chart:

```{r}
as.tibble(smry_tpr_ppr)
```


## RMSE: Root Mean Square Error
$$RMSE = \sqrt {\frac {1}{n} \sum_{i=1}^{n} (y_i - \hat y_i)^2}$$

## Appendix
`getStandingKatzMatrix` is equivalent to using the `sapply` function with the internal function `.z.HallYarborough` (the **dot** means it's internal), which we call adding the prefix `zFactor:::`. That is, the package name and **three** colons.

```{r}
# test HY with 1st-derivative using the values from the paper 
 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.3, 1.5, 1.7, 2) 
 
hy <- sapply(ppr, function(x)  
    sapply(tpr, function(y) zFactor:::.z.HallYarborough(pres.pr = x, temp.pr = y))) 
 
rownames(hy) <- tpr 
colnames(hy) <- ppr 
print(hy) 
```


## References


