---
title: "Hall-Yarborough correlation"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hall-Yarborough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=F, error=T, message=F, warning=F}
knitr::opts_chunk$set(echo=T, comment=NA, error=T, warning=F, message = F, fig.align = 'center', results="hold")
```


## How do we find the limits of accuracy in the HY correlation



## Get `z` at selected `Ppr` and `Tpr`

```{r}
# get a z value using HY
library(zFactor)

z.HallYarborough(pres.pr = 1.5, temp.pr = 2.0)
```

From the Standing-Katz chart we obtain a digitized point:

```{r}
# get a z value from the SK chart at the same Ppr and Tpr
library(zFactor)

tpr_vec <- c(2.0)
getStandingKatzMatrix(tpr_vector = tpr_vec, 
                      pprRange = "lp")[1, "1.5"]
```


## Get `z` at selected `Ppr` and `Tpr`

```{r}
library(zFactor)

z.HallYarborough(pres.pr = 1.5, temp.pr = 1.1)
```

From the Standing-Katz chart we obtain a digitized point:

```{r}
library(zFactor)

tpr_vec <- c(1.1)
getStandingKatzMatrix(tpr_vector = tpr_vec, 
                      pprRange = "lp")[1, "1.5"]
```

> We perceive a noticeable difference between the values of z from the HY calculation and the value read from the Standing-Katz chart.


## Get values of `z` for several `Ppr` and `Tpr`
In this example we provide vectors instead of a single point.

```{r}
library(zFactor)

ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.3, 1.5, 1.7, 2) 

z.HallYarborough(ppr, tpr)
```

Which is equivalent to using the sapply function with the internal function `.z.HallYarborough`, which we call adding the prefix `zFactor:::`. That is, the package name and **three** dots.

```{r}
# test HY with 1st-derivative using the values from paper 
 
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
tpr <- c(1.3, 1.5, 1.7, 2) 
 
hy <- sapply(ppr, function(x)  
    sapply(tpr, function(y) zFactor:::.z.HallYarborough(pres.pr = x, temp.pr = y))) 
 
rownames(hy) <- tpr 
colnames(hy) <- ppr 
print(hy) 
```

With the same ppr and tpr vector, we do the same for the Standing-Katz chart:

```{r}
library(zFactor)

sk <- getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr)
sk
```

Subtract and find the difference:

```{r}
err <- round((sk - hy) / sk * 100, 2)
err
```

## Error by Ppr and by PPr

```{r}
print(colSums(err))
```

```{r}
print(rowSums(err))
```

## Analyze the error for smaller values of `Tpr`

```{r}
library(zFactor)

tpr2 <- c(1.05, 1.1) 
ppr2 <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 


sk2 <- getStandingKatzMatrix(ppr_vector = ppr2, tpr_vector = tpr2, pprRange = "lp")
sk2
```

We do the same with the HY correlation:

```{r}
# calculate z values at lower values of Tpr
library(zFactor)

tpr <- c(1.05, 1.1)
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5) 

hy2 <- z.HallYarborough(pres.pr = ppr, temp.pr = tpr) 

print(hy2)
```

```{r}
err2 <- round((sk2 - hy2) / sk2 * 100, 2)
err2
```

We can see that using Hall-Yarborough correlation shows a very high error at values of `Tpr` lower or equal than `1.1` being `Tpr=1.05` the worst curve to calculate z values from.

```{r}
t_err2 <- t(err2)
t_err2
```

Applying the function `summary`:

```{r}
sum_t_err2 <- summary(t_err2)
sum_t_err2
```

We can see that the errors in `z` are considerable with a `r sum_t_err2[1,1]`% and `r sum_t_err2[6,1]`% for `Tpr=1.05`, and a `r sum_t_err2[1,2]`% and `r sum_t_err2[6,2]`% for `Tpr=1.10`


## Prepare to plot `SK` chart values vs `HY` correlation

```{r}
library(zFactor)
library(tibble)

tpr2 <- c(1.05, 1.1, 1.2, 1.3) 
ppr2 <- c(0.5, 1.0, 1.5, 2, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0, 6.5) 

sk_hy_2 <- createTidyFromMatrix(ppr2, tpr2, correlation = "HY")
as.tibble(sk_hy_2)
```

Plotting the difference between the z values in the Standing-Katz and the values calculated by Hall-Yarborough:

```{r}
library(ggplot2)

p <- ggplot(sk_hy_2, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-dif, ymax=z.calc+dif), width=.4,
                  position=position_dodge(0.05))
print(p)
```


## Analyzing the error for all the `Tpr` curves

```{r}
library(ggplot2)
library(tibble)

# get all `lp` Tpr curves
tpr_all <- getCurvesDigitized(pprRange = "lp")
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 
sk_corr_all <- createTidyFromMatrix(ppr, tpr_all, correlation = "HY")
as.tibble(sk_corr_all)


p <- ggplot(sk_corr_all, aes(x=Ppr, y=z.calc, group=Tpr, color=Tpr)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin=z.calc-dif, ymax=z.calc+dif), width=.4,
                  position=position_dodge(0.05))
print(p)
```

> The greatest errors are localized in two of the Tpr curves: at 1.05 and 1.1.



$$RMSE = \sqrt {\frac {1}{n} \sum_{i=1}^{n} (y_i - \hat y_i)^2}$$

```{r}
# MSE: Mean Squared Error
# RMSE: Root Mean Sqyared Error
# RSS: residual sum of square
# ARE:  Average Relative Error, %
# AARE: Average Absolute Relative Error, %
library(dplyr)
grouped <- group_by(sk_corr_all, Tpr, Ppr)
smry_tpr_ppr <- summarise(grouped, 
          #mean=mean(z.calc), 
          #sd=sd(z.calc), 
          RMSE= sqrt(mean((z.chart-z.calc)^2)), 
          MSE = sum((z.calc - z.chart)^2) / n(), 
          # rmse2 = sqrt(sum((z.chart-z.calc)^2)/n()),
          RSS = sum((z.calc - z.chart)^2),
          ARE = sum((z.calc - z.chart) / z.chart) * 100 / n(),
          AARE = sum( abs((z.calc - z.chart) / z.chart)) * 100 / n()
          )

ggplot(smry_tpr_ppr, aes(Ppr, Tpr)) +
           geom_tile(data=smry_tpr_ppr, aes(fill=AARE), color="white") +
    scale_fill_gradient2(low="blue", high="red", mid="yellow", na.value = "pink",
  midpoint=12.5, limit=c(0, 25), name="AARE") +
    theme(axis.text.x = element_text(angle=45, vjust=1, size=11, hjust=1))+
 coord_equal()
```




## The `Tpr` and `Ppr` values with less accuracy

```{r}
library(dplyr)

sk_hy_all %>%
    filter(Tpr %in% c("1.05", "1.1", "1.2", "1.3")) %>%
    ggplot(aes(x = z.chart, y=z.calc, group = Tpr, color = Tpr)) +
    geom_point(size = 3) +
    geom_line(aes(x = z.chart, y = z.chart), color = "black") +
    facet_grid(. ~ Tpr) +
    geom_errorbar(aes(ymin=z.calc-abs(dif), ymax=z.calc+abs(dif)), position=position_dodge(0.5), width = 0.05)
```


```{r}
library(dplyr)

sk_hy_all %>%
    filter(Tpr %in% c("1.4", "1.5", "1.6", "1.7", "1.8", "1.9")) %>%
    ggplot(aes(x = z.chart, y=z.calc, group = Tpr, color = Tpr)) +
    geom_point(size = 3) +
    geom_line(aes(x = z.chart, y = z.chart), color = "black") +
    facet_grid(. ~ Tpr) +
    geom_errorbar(aes(ymin=z.calc-abs(dif), ymax=z.calc+abs(dif)), position=position_dodge(0.5), width = 0.05)
```

```{r}
library(dplyr)

sk_hy_all %>%
    filter(as.double(Tpr) > 2 ) %>%
    ggplot(aes(x = z.chart, y=z.calc, group = Tpr, color = Tpr)) +
    geom_point(size = 3) +
    geom_line(aes(x = z.chart, y = z.chart), color = "black") +
    facet_grid(. ~ Tpr) +
geom_errorbar(aes(ymin=z.calc-abs(dif), ymax=z.calc+abs(dif)), position=position_dodge(0.5), width = 0.05)
```


```{r}
sk_hy_all
```


```{r}
sk_hy_all$z.rng <- ifelse(sk_hy_all$Ppr > 3, "zgt05", "zlt05")

q <- ggplot(sk_hy_all, aes(x=z.chart, y=z.calc)) +
    geom_line() +
    geom_point() +
     facet_grid(Tpr ~ .)
print(q)

```

```{r}

```



```{r}

q <- ggplot(sk_hy_all, aes(x=z.chart, y=z.calc, color=Tpr)) +
    geom_line() +
    geom_point() +
    facet_grid(Tpr ~ .)
print(q)

```




## Looking numerically at the errors

```{r}
# get all `lp` Tpr curves
tpr <- getCurvesDigitized(pprRange = "lp")
ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5) 

# calculate HY for the given Tpr
all_hy <- sapply(ppr, function(x)  
    sapply(tpr, function(y) z.HallYarborough(pres.pr = x, temp.pr = y))) 
rownames(all_hy) <- tpr 
colnames(all_hy) <- ppr 
cat("Calculated Hall-Yarborough\n")
print(all_hy) 

cat("\nStanding-Katz chart\n")
all_sk <- getStandingKatzMatrix(ppr_vector = ppr, tpr_vector = tpr)
all_sk

# find the error
cat("\n Errors in percentage \n")
all_err <- round((all_sk - all_hy) / all_sk * 100, 2)  # in percentage
all_err

cat("\n Errors in Ppr\n")
summary(all_err)

# for the transposed matrix
cat("\n Errors for the transposed matrix: Tpr \n")
summary(t(all_err))
```

