---
title: "Plot Standing-Katz low-pressure charts"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

If we invoke the function `getStandingKatzCurve` without any parameters, by defult will use a `Tpr=1.3` for the low-pressure set of Tpr curves and plot it.

```{r}
library(zFactor)

getStandingKatzCurve()
```

If we want to specify a Tpr curve:

```{r}
library(zFactor)

getStandingKatzCurve(tpr = 1.05, toSave = FALSE, toView = FALSE)
```

Or a couple of Tpr curves:

```{r}
library(zFactor)

tpr_vec <- c(1.05, 1.1)
result <- getStandingKatzCurve(tpr = tpr_vec, toSave = FALSE, toView = FALSE)
```

## A dataframe for each `Tpr` dataset
The returning object is a list of dataframes, which contain a dataframe for each of the Tpr curves specified:

```{r}
class(result)      # class of the object `result` is a list of dataframes
names(result)      # name of each dataframe within the list
```

For Tpr = 1.05:

```{r}
result["1.05"]
```

And, for Tpr = 1.1:
```{r}
result["1.1"]
```



```{r}
# view all the `Tpr` SK individual charts
tpr_vec <- getCurvesDigitized(pprRange = "lp")

invisible(sapply(tpr_vec, function(x) getStandingKatzCurve(tpr = x, 
                                                           toSave = FALSE, 
                                                           toView = FALSE)))
```


## All in one chart

These two functions are equivalent:

```{r}
library(zFactor)

tpr_vec <- c(1.05, 1.3, 1.5)

all_tpr1 <- (lapply(tpr_vec, function(x) getStandingKatzCurve(tpr = x, 
                                                           toSave = FALSE, 
                                                           toView = FALSE,
                                                           toPlot = FALSE)))

```


```{r}
library(zFactor)

tpr_vec <- c(1.05, 1.3, 1.5)
all_tpr2 <- (lapply(tpr_vec, function(x) getStandingKatzData(tpr = x)))
names(all_tpr2) <- tpr_vec
```


Now, we have to create a tidy datatset from the list.

```{r}
library(data.table)

all_tpr_df <- data.table::rbindlist(all_tpr2, idcol = TRUE)
colnames(all_tpr_df)[1] <- "Tpr"
```


Let's plot the curves:

```{r}
library(ggplot2)

ggplot(all_tpr_df, aes(x=Ppr, y=z, group=Tpr, color=Tpr)) + 
    geom_line() +
    geom_point()
```

The dots are digitized points of the Standing-Katz chart.

## The whole Standing-Katz chart

```{r}
library(zFactor)
library(ggplot2)
library(data.table)

tpr_vec <- c(1.05, 1.1, 1.2, 1.3, 1.5, 1.6, 1.7, 1.9, 2.0, 2.4, 2.6, 2.8, 3.0)
all_tpr2 <- (lapply(tpr_vec, function(x) getStandingKatzData(tpr = x)))
names(all_tpr2) <- tpr_vec


all_tpr_df <- data.table::rbindlist(all_tpr2, idcol = TRUE)
colnames(all_tpr_df)[1] <- "Tpr"


ggplot(all_tpr_df, aes(x=Ppr, y=z, group=Tpr, color=Tpr)) + 
    geom_line() +
    geom_point()
```

## Low range Tpr
Or we if just prefer the low range Tpr cureves:

```{r}
library(zFactor)
library(ggplot2)
library(data.table)

low_tpr_vec <- c(1.05, 1.1, 1.2, 1.3, 1.4)
low_tpr_li <- (lapply(low_tpr_vec, function(x) getStandingKatzData(tpr = x)))
names(low_tpr_li) <- low_tpr_vec


low_tpr_df <- data.table::rbindlist(low_tpr_li, idcol = TRUE)
colnames(low_tpr_df)[1] <- "Tpr"


ggplot(low_tpr_df, aes(x=Ppr, y=z, group=Tpr, color=Tpr)) + 
    geom_line() +
    geom_point()
```



```{r}
library(zFactor)
library(ggplot2)
library(data.table)

med_tpr_vec <- c(1.5, 1.6, 1.7, 1.8, 1.9)
med_tpr_li <- (lapply(med_tpr_vec, function(x) getStandingKatzData(tpr = x)))
names(med_tpr_li) <- med_tpr_vec


med_tpr_df <- data.table::rbindlist(med_tpr_li, idcol = TRUE)
colnames(med_tpr_df)[1] <- "Tpr"


ggplot(med_tpr_df, aes(x=Ppr, y=z, group=Tpr, color=Tpr)) + 
    geom_line() +
    geom_point()
```



```{r}
library(zFactor)
library(ggplot2)
library(data.table)

hi_tpr_vec <- c(2.0, 2.4, 2.6, 2.8, 3.0)
hi_tpr_li <- (lapply(hi_tpr_vec, function(x) getStandingKatzData(tpr = x)))
names(hi_tpr_li) <- hi_tpr_vec


hi_tpr_df <- data.table::rbindlist(hi_tpr_li, idcol = TRUE)
colnames(hi_tpr_df)[1] <- "Tpr"


ggplot(hi_tpr_df, aes(x=Ppr, y=z, group=Tpr, color=Tpr)) + 
    geom_line() +
    geom_point()
```


```{r}
summary(hi_tpr_df)
```


With this we are in a position to compare the original Standing-Katz chart against the most common compresssibility correlations.
