---
title: "Measuring the *error* in calculating $z$"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Practical example
To demonstrate graphically the difference between correlation and experimental data we will use the Dranchuk-AbouKassem results at a $T_{pr}$=1.05 at various $P_{pr}$.

```{r}
library(zFactor)
zFactor:::z.plot.range("DAK", interval = "fine")
```


```{r}
library(dplyr)
correlation <- "DAK"
# get all `lp` Tpr curves
tpr_all <- getStandingKatzTpr(pprRange = "lp")
# ppr <- c(0.5, 1.0, 1.5, 2.0, 2.5, 3.0, 3.5, 4.0, 4.5, 5.0, 5.5, 6.0, 6.5, 7.0)
ppr <- zFactor:::getStandingKatzPpr(interval = "fine")
sk_corr_all <- createTidyFromMatrix(ppr, tpr_all, correlation)
sk_corr_all
grouped <- group_by(sk_corr_all, Tpr, Ppr)
grouped
smry_tpr_ppr <- summarise(grouped,
                          RMSE = sqrt(mean((z.chart-z.calc)^2)),
                          MPE  = sum((z.calc - z.chart) / z.chart) * 100 / n(),
                          MAPE = sum(abs((z.calc - z.chart) / z.chart)) * 100 / n(),
                          MSE  = sum((z.calc - z.chart)^2) / n(),
                          RSS  = sum((z.calc - z.chart)^2),
                          MAE  = sum(abs(z.calc - z.chart)) / n()
)
smry_tpr_ppr
```


```{r}
library(zFactor)
library(tibble)

dak <- z.stats("DAK")
dak
dak_105 <- dak[dak$Tpr=="1.05", ]
dak_105
```

## Accuracy measurement
The comparative analysis shows tables with different error measurements:

    RMSE:  Root Mean Squared Error
    MPE:   Mean Percentage error
    MAPE:  Mean Absolute Percentage Error
    MSE:   Mean Squared Error
    RSS:   Residual sum of Squares
    MAE:   Mean Absolute Error


where:

$a_t$ are the observed true values. In our case the Standing-Katz chart $z$ values;    
$f_t$ are the calculated or predicted values (the $z$ values calculated by the correlations); and    
$n$ is the number of samples


## RMSE
Measure of accuracy, to compare errors of different calculation models for the same dataset.

$$RMSE =  \sqrt {\sum_{t=1}^n \frac {(a_t - f_t)^2} {n}}$$

<u>RMSE code</u>
```{r eval=FALSE}
RMSE = sqrt(mean((z.chart - z.calc)^2))
```



## MPE: Mean Percentage error


$$MPE = \frac {100%} {n} \sum_{t=1}^n \frac {a_t - f_t} {a_t}$$

<u>MPE code</u>
```{r eval=FALSE}
MPE  = sum((z.calc - z.chart) / z.chart) * 100 / n(),
```


## MAPE: Mean Absolute Percentage Error

$$MAPE = \frac {100} {n} \sum | \frac {a_t - f_t} {a_t}|$$

<u>MAPE code</u>
```{r eval=FALSE}
MAPE = sum(abs((z.calc - z.chart) / z.chart)) * 100 / n()
```


## MSE: Mean Squared Error
$$MSE = \frac {1}{n}  \sum_{t=1}^n (a_t - f_t)^2 $$
<u>MSE code</u>
```{r eval=FALSE}
MSE  = sum((z.calc - z.chart)^2) / n()
```



## RSS: Residual sum of Squares
$$RSS =  \sum_{t=1}^n (a_t - f_t)^2 $$
<u>RSS code</u>
```{r eval=FALSE}
RSS  = sum((z.calc - z.chart)^2)
```


## MAE: Mean Absolute Error

$$MAE = \frac {1} {n} \sum | {a_t - f_t} |$$
<u>MAE code</u>
```{r eval=FALSE}
MAE  = sum(abs(z.calc - z.chart)) / n()
```
