---
title: "R Notebook"
output: html_notebook
---

   

## Problems with using `rootSolve::uniroot.all`
We have noticed some strong deviation in he use of `rootSolve` when being applied to the solution of Hall-Yarborough correlation. It is extremely sensitive for $P_{pr}$ under 1.5.
See R chunk below.


### Calculate `z` for various values of $P_{pr}$ and $T_{pr}$

```{r results="hold"}
source("HY.R")      # modified script saved as HY.R
z.hallyarboroughL(pres.pr = 0.5, temp.pr = 1.3)$z     # SK-chart = 0.92  x
z.hallyarboroughL(pres.pr = 1.5, temp.pr = 1.3)$z     # SK-chart = 0.76  x
z.hallyarboroughL(pres.pr = 2.5, temp.pr = 1.3)$z     # SK-chart = 0.64
z.hallyarboroughL(pres.pr = 3.5, temp.pr = 1.3)$z     # SK-chart = 0.63
z.hallyarboroughL(pres.pr = 4.5, temp.pr = 1.3)$z     # SK-chart = 0.68
z.hallyarboroughL(pres.pr = 5.5, temp.pr = 1.3)$z     # SK-chart = 0.76
z.hallyarboroughL(pres.pr = 6.5, temp.pr = 1.3)$z     # SK-chart = 0.84
```

```{r}
library(rNodal)
# we use instead or temp.pr and pres.pr the values of sampling pressure and 
# temperature and gas.sg = 07 because rNodal still has not been modified to 
# take pseudo-critical or pseudo-reduced P, T.

rNodal:::z.hallyarboroughL(pres.a = 334, temp.f = 45, gas.sg = 0.7)$z
```

## Values from Standing and Katz chart
These values have been read from the SK-chart. pres.pr = 

    (0.5, 1.3) = 0.92
    (1.5, 1.3) = 0.76
    (2.5, 1.3) = 0.64
    (3.5, 1.3) = 0.63
    (4.5, 1.3) = 0.68
    (5.5, 1.3) = 0.76
    (6.5, 1.3) = 0.84
    

## Plotting z vs $P_{pr}$ at various $T_{pr}$ using `HY.R`
This cold give us the proof that using the method of finding the root for the HY equation is not appropiate to find a solution for z.

It is essential that the boundaries selection for the root finding function are properly defined. 


### Selecting a poor interval for root finding
Below the is an example of a poor selection of the interval for root finding.
In thise case, the inerval selected is `c(-5, 5.99)`.
The reason for selecting this interval is that it was purely based on two examples.
Sadly, extreme cases were not tested, such as values of $P_{pr}$ lower than 2.5 or $T_{pr}$ lower than 1.4.

We will test in the following paragraphs other intervals.
We start by makng a function out of the chunk below first.

```{r}
source("HY.R")  

ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.6)
tpr <- c(1.3, 1.5, 1.7, 2)
tbl <- sapply(ppr, function(x) 
    sapply(tpr, function(y) z.hallyarboroughL(pres.pr = x, temp.pr = y,
                                              interval = c(-5, 5.99))$z))
tbl

library(ggplot2)
plot(x = ppr, y = tbl[1,], type = "l", main = "z @ Tpr = 1.3")
plot(x = ppr, y = tbl[2,], type = "l", main = "z @ Tpr = 1.5")
plot(x = ppr, y = tbl[3,], type = "l", main = "z @ Tpr = 1.7")
plot(x = ppr, y = tbl[4,], type = "l", main = "z @ Tpr = 2.0")
```

## Testing other intervals for the root finding



```{r}
source("HY.R")  
library(ggplot2)

testIntervals <- function(interval) {
    ppr <- c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.6)
    tpr <- c(1.3, 1.5, 1.7, 2)
    tbl <- sapply(ppr, function(x) 
        sapply(tpr, function(y) z.hallyarboroughL(pres.pr = x, temp.pr = y,
                                                  interval = interval)$z))
    print(tbl)
    
    plot(x = ppr, y = tbl[1,], type = "l", main = "z @ Tpr = 1.3")
    plot(x = ppr, y = tbl[2,], type = "l", main = "z @ Tpr = 1.5")
    plot(x = ppr, y = tbl[3,], type = "l", main = "z @ Tpr = 1.7")
    plot(x = ppr, y = tbl[4,], type = "l", main = "z @ Tpr = 2.0")
}
```

```{r}
testIntervals(c(0, 1))
```

```{r}
testIntervals(c(0, 2))
```


### Discontinuity at Upper Limit (UL) = 100
It is safe to navigate the range 0 to 99.99. But solving the equation will fail at exactly UL = 100 because of inifinite values generated. All values above 100 will yield zero on all Ppr and Tpr supplied.


```{r}
testIntervals(c(0, 101))
```

Upper bound values above 4.3 also brings some discontinuity when the lower bound is, for instance, -1.

```{r}
testIntervals(c(-1, 4.4))
```

Or the interval (-9, 5).

```{r}
testIntervals(c(-9, 5))
```

The interval (-20, 20) seems safe but (-20, 100) is not.

```{r}
testIntervals(c(-20, 20))
```

